Certified Knowledge Compilation \\ with Application to Verified Model Counting \\ Experimental Details

### 2023-08-27 doc/JAIR/jair-data.tex
Include one-sided data
Completed overall data

     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 122 instances, Min = 304, Median = 774,883, Mean = 72,389,514, Max = 1,765,743,261
       \end{itemize}

@@

     \item Performance (122 instances):
       \begin{itemize}
       \item Times: Min = 0.012 secs, Max = 2804.5 secs, Median = 3.46 secs
       \item Defining clauses: Min = 304, Max = 1,765,743,261, Median = 774,883
       \item Tree Sizes ($D_t$): Min = 304, max = 14,924,670,455,484, median = 7,669,582.
       \item Tree ratio: Min = 1.0, Max = 52,409.7, Median = 11.61, Harmonic Mean = 3.21
       \end{itemize}
     \end{itemize}
\end{enumerate}

@@

General findings:
\begin{itemize}
\item The tree ratio provides a reasonable discriminator for monolithic vs.~structured.  Based on experiments, set threshold at 5.0. Use monolithic below this and structured above.
\item We could not find any way to exploit the hybrid approach for these benchmarks.
\item On the other hand, the monolithic approach becomes impractical beyond around $10^7$ defining clauses.  Even the best SAT solvers become unstable beyond this.
\item Therefore, use hybrid approach for very large problems with tree ratios below 5.0, shift to monolithic once the tree size is at most $10^7$.
\end{itemize}

@@

\begin{itemize}
\item Of the 122 benchmarks, 100 were completed using the hybrid approach within 10,000 seconds.
\end{itemize}

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.30,0.96)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.4,
                              xmode=log,
                              xmin=0.01,xmax=4000,
                              xtick={0.01, 0.1,1.0,10,100,1000,10000}, xticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}, {10,000}},
                              ymin=0.01, ymax=29000,
                              ytick={0.01, 0.1,1.0,10,100,1000,10000,100000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000},{10,000},{100,000}},
                              xlabel={D4 runtime (seconds)}, ylabel={CPOG generation and checking runtime (seconds)},
%                              title={D4 Defining Clause Generation}
            ]
    \input{data-formatted/seconds-summary}
%    \input{data-formatted/seconds-summary-onesided}
    \input{data-formatted/seconds-summary-failures}
    \legend{
      \scriptsize \textsf{Full validation},
%      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{Resource limit exceeded},
    }
    \addplot[mark=none, dashed] coordinates{(0.01,25000) (4000, 25000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,20000) (4000,20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (2000,20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (200, 20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (20, 20000)};
    \node[left] at (axis cs: 0.9,0.12) {$0{.}1\times$};
    \node[left] at (axis cs: 0.09,0.12) {$1\times$};
    \node[left] at (axis cs: 0.09,1.2) {$10\times$};
    \node[left] at (axis cs: 0.09,12.0) {$100\times$};
    \node[left] at (axis cs: 0.09,120.0) {$1000\times$};
       \end{axis}

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10},
      legend style={at={(0.92,0.28)}},
      legend cell align={left},
                              x post scale=1.9, y post scale=2.10,
                              xmode=log,xmin=100,xmax=1e10, 
                              xtick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              xticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              ymode=log, ymin=100, ymax=1800000000, 
                              ytick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000},
                              yticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$},
                              xlabel={Defining Clauses}, ylabel={Proof Clauses},
            ]
    \input{data-formatted/clauses-summary}
%%    \input{data-formatted/clauses-defining-onesided}
    \input{data-formatted/clauses-summary-failures}

    \legend{
      \scriptsize \textsf{Full validation},
%      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{No validation},
    }
    \addplot[mark=none, color=lightblue] coordinates{(100,100) (1e9,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1000) (1e8,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,10000) (1e7,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,100000) (1e6,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1000000) (1e5,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1e9) (1e10,1e9)};
    \addplot[color=black,dashed] coordinates{(1e2,1.5e9) (1e10,1.5e9)};
    \node[left] at (axis cs: 900,1200) {$1\times$};
    \node[left] at (axis cs: 900,12000) {$10\times$};
    \node[left] at (axis cs: 900,120000) {$100\times$};
    \node[left] at (axis cs: 900,1200000) {$1{,}000\times$};
    \node[left] at (axis cs: 900,12000000) {$10{,}000\times$};

>>>>

     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 123 instances, Min = 304, Median = 774,883, Mean = 72,389,514, Max = 2,761,457,765 (1,765,743,261 for those that didn't overflow).
       \end{itemize}

@@

     \item Performance (122 instances.  Don't include the one that overflowed):
       \begin{itemize}
       \item Times: Min = 0.012 secs, Max = 2804.5 secs, Median = 3.46 secs
       \item Defining clauses: Min = 304, Max = 1,765,743,261, Median = 774,883
       \item Tree Sizes ($D_t$): Min = 304, max = 14,924,670,455,484, median = 7,669,582.
       \item Tree ratio: Min = 1.0, Max = 52,409.7, Median = 11.61, Harmonic Mean = 3.21
       \end{itemize}
     \end{itemize}
\end{enumerate}

@@

General findings:
\begin{itemize}
\item The tree ratio provides a reasonable discriminator for monolithic vs.~structured.  Based on experiments, set threshold at 5.0. Use monolithic below this and structured above.
\item We could not find any way to exploit the hybrid approach for the smaller benchmarks.
\item On the other hand, the monolithic approach becomes impractical beyond around $10^6$ defining clauses.  Even the best SAT solvers become unstable beyond this.
\item Therefore, use hybrid approach for very large problems with tree ratios below 5.0, shift to monolithic once the tree size is at most $10^6$.
\item Also tried with limit of $10^5$ clauses.  Generally did worse in terms of completions.
\end{itemize}

@@

\begin{itemize}
\item Of the 123 benchmarks, 110 were completed using the hybrid approach within 10,000 seconds.
\item One additional one completed within 10,000 for structural (and for hybrid with a threshold of $10^5$ clauses.
\item For seven others, was able to do one-sided proof
\item Only one had more than $10^9$ defining clauses.
\item Total of 5 (including 1 overflow case) for which no form of validation succeeded.
  \begin{itemize}
  \item Three of these had more than $10^9$ defining clauses
  \item Smallest had 405 million

  \end{itemize}
\end{itemize}

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.30,0.96)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.4,
                              xmode=log,
                              xmin=0.01,xmax=4000,
                              xtick={0.01, 0.1,1.0,10,100,1000,10000}, xticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}, {10,000}},
                              ymin=0.01, ymax=29000,
                              ytick={0.01, 0.1,1.0,10,100,1000,10000,100000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000},{10,000},{100,000}},
                              xlabel={D4 runtime (seconds)}, ylabel={CPOG generation and checking runtime (seconds)},
%                              title={D4 Defining Clause Generation}
            ]
    \input{data-formatted/seconds-summary}
    \input{data-formatted/seconds-summary-structured-only}
    \input{data-formatted/seconds-summary-onesided}
    \input{data-formatted/seconds-summary-failures}
    \legend{
      \scriptsize \textsf{Full validation, Hybrid},
      \scriptsize \textsf{Full validation, Structural},
      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{Resource limit exceeded},
    }
    \addplot[mark=none, dashed] coordinates{(0.01,25000) (4000, 25000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,20000) (4000,20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (2000,20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (200, 20000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,10.0) (20, 20000)};
    \node[left] at (axis cs: 0.9,0.12) {$0{.}1\times$};
    \node[left] at (axis cs: 0.09,0.12) {$1\times$};
    \node[left] at (axis cs: 0.09,1.2) {$10\times$};
    \node[left] at (axis cs: 0.09,12.0) {$100\times$};
    \node[left] at (axis cs: 0.09,120.0) {$1000\times$};
       \end{axis}

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10},
      legend style={at={(0.92,0.28)}},
      legend cell align={left},
                              x post scale=1.9, y post scale=2.10,
                              xmode=log,xmin=100,xmax=1e10, 
                              xtick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              xticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              ymode=log, ymin=100, ymax=1800000000, 
                              ytick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000},
                              yticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$},
                              xlabel={Defining Clauses}, ylabel={Proof Clauses},
            ]
    \input{data-formatted/clauses-summary}
    \input{data-formatted/clauses-summary-structured-only}
    \input{data-formatted/clauses-summary-onesided}
    \input{data-formatted/clauses-summary-failures}

    \legend{
      \scriptsize \textsf{Full validation, Hybrid},
      \scriptsize \textsf{Full validation, Structural},
      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{No validation},
    }
    \addplot[mark=none, color=lightblue] coordinates{(100,100) (1e9,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1000) (1e8,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,10000) (1e7,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,100000) (1e6,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1000000) (1e5,1e9)};
    \addplot[mark=none, color=lightblue] coordinates{(100,1e9) (1e10,1e9)};
    \addplot[color=black,dashed] coordinates{(1e2,1.5e9) (1e10,1.5e9)};
    \node[left] at (axis cs: 900,1200) {$1\times$};
    \node[left] at (axis cs: 900,12000) {$10\times$};
    \node[left] at (axis cs: 900,120000) {$100\times$};
    \node[left] at (axis cs: 900,1200000) {$1{,}000\times$};
    \node[left] at (axis cs: 900,12000000) {$10{,}000\times$};
